---
title: "sediment - mecanistic"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "Sediment - ACP"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Loadings

## Load packages

```{r packages, message = FALSE}
library(tidyverse)
library(here)

library(readxl)
library(readr) # write .csv
library(parzer) # read GPS coordinates

library(nlme)
library(lmerTest)

# ACP
library("FactoMineR")
library("factoextra")
library("missMDA")

# graphics
library(ggplot2)
library(gridExtra)
library(RColorBrewer)
```

## Load contaminant lists and labels
```{r}
sed_contam <- read_csv(file = here("data/sed_contam.csv"))
```



#--------------
# Create a table of data for sediment:

sediment = select(.data = sed_contam,
                      lat, lon,
                      site,
                      year,
                      zone_publi,
                      season,
                      `X<63`,
                      `Ctot_mg_g-1`,
                      `C/N_mol`,
                      "a-HBCDD":"HFPO-DA")

sediment$year = as.factor(sediment$year)

# get coordinates in a usable format for graphics
sediment$lat = parzer::parse_lat(lat = sediment$lat)
sediment$lon = parzer::parse_lon(lon = sediment$lon)

# add an upstream point for dist reference
sediment = mutate(sediment,
                    dist = round(distGeo(p1 = cbind(lon, lat),
                                         p2 = c(0.28, 49.437))/1000, digits = 1))

# translate seasons
sediment$season = factor(sediment$season, 
                           labels = c("Spring",
                                      "Autumn"),
                           levels = c("Printemps",
                                      "Automne"))

# add a campaign column
sediment = mutate(sediment,
                    campaign = paste(season, year))

sediment_PCB = select(.data = sediment,
                      -c("PFPeA":"HFPO-DA"),
                      -c("a-HBCDD":"g-HBCDD"))

sediment_PFAS = select(.data = sediment,
                      -c("CB28":"CB194"),
                      -c("a-HBCDD":"g-HBCDD"))

sediment_HBCDD = select(.data = sediment,
                       -c("CB28":"CB194"),
                       -c("PFPeA":"HFPO-DA"))



# change table structure for grain size: add a column "grain size" with factor values = `X<63`:`1000<X<2000`
sediment_L = pivot_longer(sediment,
                          cols = "CB28":"HFPO-DA",
                          names_to = "Contam_TAG",
                          values_to = "Contam_ng.gdw")

sediment_L$Contam_TAG = as.factor(sediment_L$Contam_TAG)

sediment_PCB_L = pivot_longer(sediment_PCB,
                          cols = "CB28":"CB194",
                          names_to = "Contam_TAG",
                          values_to = "Contam_ng.gdw")

sediment_PCB_L$Contam_TAG = as.factor(sediment_PCB_L$Contam_TAG)

#-----------------------

# res.comp = imputeFAMD(ACP_benthos, ncp=5)
# res.famd = FAMD(res.comp$completeObs)
res.comp = imputeFAMD(X = as.data.frame(sediment_PCB_L[,-c(1:4)]), ncp = 3)
res.famd = FAMD(res.comp$completeObs)

summary(res.famd)


fviz_screeplot(res.pca)

fviz_pca_var(res.pca, col.var = "contrib", 
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel=T
)
fviz_pca_var(res.pca, col.var = "contrib", 
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel=T, axes = c(2,3)
)

#-----------------------
# Fine-sand grain portion => *** ###
glm_sed_grain <- glm(Contam_ng.gdw ~ `X<63`, data = sediment_L)
summary(glm_sed_grain) # 4.92e-10 ***
glm_sed_grain_PCB <- glm(Contam_ng.gdw ~ `X<63`, data = sediment_PCB_L)
summary(glm_sed_grain_PCB) # 3.3e-16 ***
ggplot(sediment_PCB_L) +
  aes(y=Contam_ng.gdw, x= `X<63`, col=Contam_TAG) +
  geom_point() +
  geom_smooth(method = "lm", alpha=0.1)


#-----------------------
# TOC => 0.24 
glm_sed_TOC <- glm(Contam_ng.gdw ~ `Ctot_mg_g-1`, data = sediment_L)
summary(glm_sed_TOC)
glm_sed_TOC_PCB <- glm(Contam_ng.gdw ~ `Ctot_mg_g-1`, data = sediment_PCB_L)
summary(glm_sed_TOC_PCB) # 3.3e-16 ***
ggplot(sediment_PCB_L) +
  aes(y=Contam_ng.gdw, x= `Ctot_mg_g-1`, col=Contam_TAG) +
  geom_point() +
  geom_smooth(method = "lm", alpha=0.1)

#-----------------------
# C/N => 0.0802
glm_sed_CoverN <- glm(Contam_ng.gdw ~ `C/N_mol`, data = sediment_L)
summary(glm_sed_CoverN)
glm_sed_CoverN_PCB <- glm(Contam_ng.gdw ~ `C/N_mol`, data = sediment_PCB_L)
summary(glm_sed_CoverN_PCB) # 3.3e-16 ***
ggplot(sediment_PCB_L) +
  aes(y=Contam_ng.gdw, x= `C/N_mol`, col=Contam_TAG) +
  geom_point() +
  geom_smooth(method = "lm", alpha=0.1)

#-----------------------
# Season => 0.00186 ** ###
glm_sed_season <- glm(Contam_ng.gdw ~ season, data = sediment_L)
summary(glm_sed_season)
glm_sed_season_PCB <- glm(Contam_ng.gdw ~ `season`, data = sediment_PCB_L)
summary(glm_sed_season_PCB) # 3.3e-16 ***
ggplot(sediment_PCB_L) +
  aes(y=Contam_ng.gdw, x= `season`, col=Contam_TAG) +
  geom_point() +
  geom_smooth(method = "lm", alpha=0.1)

#-----------------------
# Distance to upstream => 0.409
glm_sed_dist <- glm(Contam_ng.gdw ~ dist, data = sediment_L)
summary(glm_sed_dist)
glm_sed_dist_PCB <- glm(Contam_ng.gdw ~ `dist`, data = sediment_PCB_L)
summary(glm_sed_dist_PCB) # 3.3e-16 ***
ggplot(sediment_PCB_L) +
  aes(y=Contam_ng.gdw, x= `dist`, col=Contam_TAG) +
  geom_point() +
  geom_smooth(method = "lm", alpha=0.1)

#-----------------------
# Bank => 0.00175 ** ###
glm_sed_bank <- glm(Contam_ng.gdw ~ zone_publi, data = sediment_L)
summary(glm_sed_bank)
glm_sed_bank_PCB <- glm(Contam_ng.gdw ~ `zone_publi`, data = sediment_PCB_L)
summary(glm_sed_bank_PCB) # 3.3e-16 ***
ggplot(sediment_PCB_L) +
  aes(y=Contam_ng.gdw, x= `zone_publi`, col=Contam_TAG) +
  geom_point() +
  geom_smooth(method = "lm", alpha=0.1)


#############################

ggplot(sediment_L) +
  aes(x = `X<63`,
      y = Contam_ng_g.1Corg,
      col = Contam_TAG) +
  geom_smooth()

##########################
# random effect on interecpt only

gD=groupedData(data = sediment_L, Contam_ng_g.1Corg~`X<63`|Contam_TAG,
               labels=list(x="X<63",y="Contamination"),
               units=list(x="%",y="ng_g.1Corg"))

plot(gD, outer=~1, key = list(lines=F), strip = F, aspect = 1)

sed_L = na.omit(sediment_L)
sed_L = rename(sed_L, fine_grain = `X<63`)

lme_tmp = lme(Contam_ng_g.1Corg ~ fine_grain, random = ~ 1|Contam_TAG, data = sed_L)
summary(lme_tmp)
plot(lme_tmp)

lme_tmp2 = lme(Contam_ng_g.1Corg ~ fine_grain + season, random = ~ 1|Contam_TAG, data = sed_L)
summary(lme_tmp2)
plot(lme_tmp2)

lme_tmp3 = lme(Contam_ng_g.1Corg ~ fine_grain + season + zone_publi, random = ~ 1|Contam_TAG, data = sed_L)
summary(lme_tmp3)
plot(lme_tmp3)

pdSymm(

##########################
# random effect on interecept and slopes
lmer_tmp = lmer(Contam_ng_g.1Corg ~ fine_grain + (1|Contam_TAG), data = sed_L)
summary(lmer_tmp)

lmer_tmp2 = lmer(Contam_ng_g.1Corg ~ fine_grain + season + (season|Contam_TAG), data = sed_L)
summary(lmer_tmp2)

lmer_tmp3 = lmer(Contam_ng_g.1Corg ~ fine_grain + season + zone_publi + 
                   (1+season|Contam_TAG) + (1+zone_publi|Contam_TAG), data = sed_L)
summary(lmer_tmp3)


##########################
glm_sed <- glm(Contam_ng_g.1Corg ~ `X<63` + season + zone_publi, data = sediment_L)
summary(glm_sed)
plot(glm_sed)


lme_sed <- lmer(Contam_ng_g.1Corg ~ `X<63` + season + zone_publi +
                 (1 + `X<63` | Contam_TAG) +
                 (1 + `season` | Contam_TAG) +
                 (1 + `zone_publi` | Contam_TAG),
                 data = sediment_L)
summ_lme_sed <-  summary(lme_sed)




plot(benthos_contam$lip_PS_percent, benthos_contam$sommePCB_ng_gdw,
     ylab = "Somme des PCB (ng.g-1 ps)",
     xlab = "Taux de lipides (% ps)")
lm1 = lm(benthos_contam$sommePCB_ng_gdw ~ benthos_contam$lip_PS_percent)
if(summary(lm1)$coef[2,4]<0.05){
  abline(lm1)
}





plot(benthos_contam$lip_PS_percent, benthos_contam$sommeHBCDD_ng_gdw,
     ylab = "Somme des HBCDD (ng.g-1 ps)",
     xlab = "Taux de lipides (% ps)")
lm2 = lm(benthos_contam$sommeHBCDD_ng_gdw ~benthos_contam$lip_PS_percent)
summary(lm2)
if(summary(lm2)$coef[2,4]<0.05){
  abline(lm2)
}
