---
title: "Contenus stomacaux"
author: "Florence Mounier"
date: "2025-01-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(chopin.project)
library("RColorBrewer")
library(tidyverse)
library(gridExtra)
library(scales)
```

## Statistiques descriptives

```{r}
summary(as.factor(stomac_soles$Statut))
summary(as.factor(stomac_soles$N.poissons))
summary(as.factor(stomac_soles$espèce.stade))
summary(as.factor(stomac_soles$secteur))
```

```{r}
levels(stomac_soles$`espèce.stade`)
```

Prey species names:
```{r}
levels(stomac_soles$ScientificName_accepted)
ordered_alphab_preys = table(stomac_soles$ScientificName_accepted)
ordered_occur_preys = summary(stomac_soles$ScientificName_accepted)
```
Number of fish:
```{r}
length(levels(stomac_soles$`N.poissons`))
```
Sectors:
```{r}
levels(stomac_soles$secteur)
```
List of subdatasets of stomac content of interest:

"Chenal", "Embouchure", "Fosse Nord", "Fosse Sud", "Sole G0", 
"Sole G0 printemps", "Sole G0 automne", "G0_CH", "G0_EM", "G0_FN", 
"G0_FS", "Sole G1", "Sole G1 printemps", "Sole G1 automne", "G1_CH", 
"G1_EM", "G1_FN", "G1_FS", "Sole G2", "G2_EM", "G2_FN"
```{r}
list_grps = list_grp(stomac_soles)
```


## Stomach content for different sub datasets of soles

Pie charts + table of values
```{r}
min_abundance = 0.1
min_occurrence = 0.1

res_all_soles = get_res_diet(
  main = "ALL SOLES",
  grp = stomac_soles,
  min_abond = min_abundance,  min_occur = min_occurrence
)

res_G0 = get_res_diet(
  main = "Sole G0",
  grp = stomac_soles[which(stomac_soles$espèce.stade == "Sole G0"), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)
res_G1 = get_res_diet(
  main = "Sole G1",
  grp = stomac_soles[which(stomac_soles$espèce.stade == "Sole G1"), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)
res_G2 = get_res_diet(
  main = "Sole G2",
  grp = stomac_soles[which(stomac_soles$espèce.stade == "Sole G2"), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)

res_G0_juin = get_res_diet(
  main = "Sole G0 printemps",
  grp = stomac_soles[which(stomac_soles$espèce.stade == "Sole G0" &
                             stomac_soles$Campagne == "Print-17"), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)
res_G0_oct = get_res_diet(
  main = "Sole G0 automne",
  grp = stomac_soles[which(stomac_soles$espèce.stade == "Sole G0" &
                             stomac_soles$Campagne == "Aut-17"), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)

res_G1_juin = get_res_diet(
  main = "Sole G1 printemps",
  grp = stomac_soles[which(stomac_soles$espèce.stade == "Sole G1" &
                             stomac_soles$Campagne == "Print-17"), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)
res_G1_oct = get_res_diet(
  main = "Sole G1 automne",
  grp = stomac_soles[which(stomac_soles$espèce.stade == "Sole G1" &
                             stomac_soles$Campagne == "Aut-17"), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)


res_che = get_res_diet(
  main = "Chenal",
  grp = stomac_soles[which(stomac_soles$secteur == "Chenal"), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)
res_G0_che = get_res_diet(
  main = "G0 Chenal",
  grp = stomac_soles[which(stomac_soles$secteur == "Chenal" &
                             stomac_soles$espèce.stade ==
                             "Sole G0"), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)
res_G0_che_juin = get_res_diet(
  main = "G0 Chenal printemps",
  grp = stomac_soles[which(
    stomac_soles$espèce.stade == "Sole G0" &
      stomac_soles$espèce.stade == "Sole G0" &
      stomac_soles$Campagne == "Print-17"
  ), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)
res_G0_che_oct = get_res_diet(
  main = "G0 Chenal automne",
  grp = stomac_soles[which(
    stomac_soles$espèce.stade == "Sole G0" &
      stomac_soles$espèce.stade == "Sole G0" &
      stomac_soles$Campagne == "Aut-17"
  ), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)


res_emb = get_res_diet(
  main = "Embouchure",
  grp = stomac_soles[which(stomac_soles$secteur == "Embouchure"), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)
res_G0_emb = get_res_diet(
  main = "G0 Embouchure",
  grp = stomac_soles[which(stomac_soles$secteur == "Embouchure" &
                             stomac_soles$espèce.stade == "Sole G0"), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)
res_G0_emb_juin = get_res_diet(
  main = "G0 Embouchure printemps",
  grp = stomac_soles[which(
    stomac_soles$secteur == "Embouchure" &
      stomac_soles$espèce.stade == "Sole G0" &
      stomac_soles$Campagne == "Print-17"
  ), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)
res_G0_emb_oct = get_res_diet(
  main = "G0 Embouchure automne",
  grp = stomac_soles[which(
    stomac_soles$secteur == "Embouchure" &
      stomac_soles$espèce.stade == "Sole G0" &
      stomac_soles$Campagne == "Aut-17"
  ), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)


res_FS = get_res_diet(
  main = "Fosse Sud",
  grp = stomac_soles[which(stomac_soles$secteur == "Fosse Sud"), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)
res_G0_FS = get_res_diet(
  main = "G0 Fosse Sud",
  grp = stomac_soles[which(stomac_soles$secteur == "Fosse Sud" &
                             stomac_soles$espèce.stade == "Sole G0"), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)
res_G0_FS_juin = get_res_diet(
  main = "G0 Fosse Sud printemps",
  grp = stomac_soles[which(
    stomac_soles$secteur == "Fosse Sud" &
      stomac_soles$espèce.stade == "Sole G0" &
      stomac_soles$Campagne == "Print-17"
  ), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)
res_G0_FS_oct = get_res_diet(
  main = "G0 Fosse Sud automne",
  grp = stomac_soles[which(
    stomac_soles$secteur == "Fosse Sud" &
      stomac_soles$espèce.stade == "Sole G0" &
      stomac_soles$Campagne == "Aut-17"
  ), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)


res_FN = get_res_diet(
  main = "Fosse Nord",
  grp = stomac_soles[which(stomac_soles$secteur == "Fosse Nord"), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)
res_G0_FN = get_res_diet(
  main = "G0 Fosse Nord",
  grp = stomac_soles[which(stomac_soles$secteur == "Fosse Nord" &
                             stomac_soles$espèce.stade == "Sole G0"), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)
res_G0_FN_juin = get_res_diet(
  main = "G0 Fosse Nord printemps",
  grp = stomac_soles[which(
    stomac_soles$secteur == "Fosse Nord" &
      stomac_soles$espèce.stade == "Sole G0" &
      stomac_soles$Campagne == "Print-17"
  ), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)
res_G0_FN_oct = get_res_diet(
  main = "G0 Fosse Nord automne",
  grp = stomac_soles[which(
    stomac_soles$secteur == "Fosse Nord" &
      stomac_soles$espèce.stade == "Sole G0" &
      stomac_soles$Campagne == "Aut-17"
  ), ],
  min_abond = min_abundance,  min_occur = min_occurrence
)

ALL_res = list(res_G0_che_juin$`med %abun`, res_G0_che_oct$`med %abun`,
               res_G0_emb_juin$`med %abun`, res_G0_emb_oct$`med %abun`,
               res_G0_FS_juin$`med %abun`, res_G0_FS_oct$`med %abun`,
               res_G0_FN_juin$`med %abun`, res_G0_FN_oct$`med %abun`,
               res_G1_juin$`med %abun`, res_G1_oct$`med %abun`,
               res_G2$`med %abun`)

```

Change colors: 1 color for each species with similar colors per taxon

```{r}
list_species = c()
for(t in 1:length(ALL_res)){
  list_species = c(list_species, as.character(ALL_res[[t]]$species))
}
especes = levels(as.factor(list_species))

list_taxon = c()
for(s in 1:length(especes)){
  list_taxon = c(list_taxon, as.character(table_species$taxon[which(table_species$species==especes[s])]))
}
list_taxon = data.frame("species"=especes, "taxon" = list_taxon)
summary(list_taxon)

col_annelid1 = brewer.pal(n = 9, name = 'OrRd')[c(1,2,3,4,6,7,8,9)]
col_annelid2 = brewer.pal(n = 9, name = 'YlOrRd')[c(1,3,4,6,7,8,9)]
col_arthro1 = brewer.pal(n = 9, name = 'Greens')[-1]
col_arthro2 = brewer.pal(n = 8, name = 'Dark2')[c(1,3,5,8)]
col_mollusc1 = brewer.pal(n = 9, name = 'Purples')[c(3,5,7,8,9)]
col_mollusc2 = brewer.pal(n = 9, name = 'Blues')[c(4,6,8,9)]

list_taxon = list_taxon[order(list_taxon$taxon),]
list_taxon$col_species = c(col_annelid1,col_annelid2,
                           col_arthro1,col_arthro2, "grey", col_mollusc1,col_mollusc2)
```


```{r}
b1=piechart_diet(res_G0_emb_juin$`med %abun`)+ ggtitle("G0 embouchure printemps")
b3=piechart_diet(res_G0_che_juin$`med %abun`)+ ggtitle("G0 chenal printemps")
b5=piechart_diet(res_G0_FN_juin$`med %abun`)+ ggtitle("G0 Fosse Nord printemps")
b7=piechart_diet(res_G0_FS_juin$`med %abun`)+ ggtitle("G0 Fosse Sud printemps")

b2=piechart_diet(res_G0_emb_oct$`med %abun`)+ ggtitle("G0 embouchure automne")
b4=piechart_diet(res_G0_che_oct$`med %abun`)+ ggtitle("G0 chenal automne")
b6=piechart_diet(res_G0_FN_oct$`med %abun`)+ ggtitle("G0 Fosse Nord automne")
b8=piechart_diet(res_G0_FS_oct$`med %abun`)+ ggtitle("G0 Fosse Sud automne")

b9=piechart_diet(res_G1_juin$`med %abun`)+ ggtitle("G1 printemps")
b10=piechart_diet(res_G1_oct$`med %abun`)+ ggtitle("G1 automne")

b11=piechart_diet(res_pie=res_G2$`med %abun`)+ ggtitle("G2 automne")

png("Contenus_stomacaux_classes_age.jpeg",
    width = 25, height = 50, units = "cm",res=720)
grid.arrange(b1,b2,b3,b4,b5,b6,b7,b8,b9,b10,b11,ncol=2,nrow=6,
             layout_matrix = matrix(c(seq(1,10,by = 1),NA,11),ncol = 2, nrow=6,byrow = T))
dev.off()
```


```{r}
p1=piechart_diet(res_G0_che$`med %abun`) + ggtitle("Soles G0 Chenal")
p2=piechart_diet(res_G0_emb$`med %abun`) + ggtitle("Soles G0 Embouchure")
p3=piechart_diet(res_G0_FS$`med %abun`) + ggtitle("Soles G0 Fosse Sud")
p4=piechart_diet(res_G0_FN$`med %abun`) + ggtitle("Soles G0 Fosse Nord")
b2=piechart_diet(res_G1$`med %abun`)+ ggtitle("Tous les G1")
b3=piechart_diet(res_G2$`med %abun`)+ ggtitle("Tous les G2")

png("Contenus_stomacaux.jpeg",width = 30, height = 25, units = "cm",res=720)
grid.arrange(p1,p2,p3,p4,b2,b3,ncol=2,nrow=3)
dev.off()
```

